---
title: "Heart Disease Analysis"
author: "Dr. Shashikanth Vydyula"
date: '`r Sys.Date()`'
output:
  html_document:
    df_print: paged
    toc: true
    number_sections: false
    highlight: pygments
    code_folding: hide
---

<style type="text/css">

h1.title {
  font-size: 38px;
  color: DarkRed;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
</style>


***

```{r steup, include=FALSE}
knitr::opts_chunk$set(
  comment = NA,
  warning= FALSE,
  message = FALSE
)
```

# Load Packages
```{r}
library('tidyverse') # essential R packages for data science
library('data.table')

library('ggplot2') # visualization
library("ggtext")
library("highcharter")

library('dplyr') # data manipulation
library('readr') # input/output
library('skimr') # overview
library('tibble') # data wrangling
library('tidyr') # data wrangling
library('purrr') # data wrangling
library('stringr') # string manipulation
library('forcats') # factor manipulation
library('janitor') # data cleaning
```

# Import Data
```{r}
data<-read.csv("C:/Shashi/Data & Dashboards/Datasets/heart.csv")

summary(data)
str(data)
```

***


> **Before manipulating the data, let us document the values in the data and what they mean while we manipulate along.**

```{r}
#1. CP or Chest Pain
#**cp (Chest Pain) : Chest Pain type chest pain type**
#**Value 0: typical angina**
#**Value 1: atypical angina**
#**Value 2: non-anginal **
#**Value 3: asymptomatic**

data2<-data %>% mutate(cp = case_when(cp == 0 ~ 'TYPICAL ANGINA',
                                        cp == 1 ~ 'ATYPICAL ANGINA',
                                        cp ==2  ~ 'NON ANGINAL',
                                        cp ==3 ~ 'ASYMPTOMATIC')) %>% 
#Sex (1=Male, 0=Female)
 mutate(sex=case_when(sex == 0 ~ "Female",
                                     sex == 1 ~ "Male")) %>% 

#fbs : (fasting blood sugar > 120 mg/dl) (1 = true; 0 = false)
  mutate(fbs=case_when(fbs == 0 ~ "<120 mg/dl",
                                     fbs ==1 ~ ">120 mg/dl")) %>% 


#rest_ecg : resting electrocardiographic results
#**Value 0: normal**
#**Value 1: having ST-T wave abnormality (T wave inversions and/or ST elevation or depression of > 0.05 mV)**
#**Value 2: showing probable or definite left ventricular hypertrophy by Estes' criteria**

mutate(restecg=case_when(restecg==0 ~ "Normal",restecg==1 ~ "ST-T Wave Abnormality", restecg==2 ~ "Probable or Definite Left Ventricular Hypertrophy")) %>% 


#exang: exercise induced angina (1 = yes; 0 = no)
 mutate(exng=case_when(exng==0 ~ "No",
                                      exng==1 ~ "Yes")) %>% 

#output 1=Heart Disease, 0=Normal
 mutate(output=case_when(output==0 ~ "Normal",
                                        output==1 ~ "Heart Disease"))
```

***


# Presence of heart Disease
```{r}
heart_Disease<-data2 %>% 
  group_by(output) %>% 
  summarise(count=n()) %>% 
  mutate(perc=paste0(round(count / sum(count) * 100, 1), "%")) 
  
view(heart_Disease)

highchart() %>% 
  hc_chart(type ="column",
           options3d = list(enabled = TRUE, alpha = 15, beta = 15)) %>%
  hc_xAxis(categories = heart_Disease$output) %>% 
  hc_add_series(data = heart_Disease$count, name = "Heart Disease/Normal Count") %>%
  hc_add_theme(hc_theme_economist()) %>%
  hc_title(text="Heart Disease Incidence")
```

***

# Blood sugar
```{r}
fbs<-data2 %>% 
  group_by(fbs) %>% 
  summarise(count=n()) %>% 
  mutate(fbs_perc=paste0(round(count/sum(count)*100, 2),"%"))

view(fbs)

highchart() %>% 
  hc_chart(type ="column",
           options3d = list(enabled = TRUE, alpha = 15, beta = 15)) %>%
  hc_xAxis(categories = fbs$fbs) %>% 
  hc_add_series(data = fbs$count, name = "Bloog Sugar Levels - Fasting") %>%
  hc_add_theme(hc_theme_monokai()) %>%
  hc_title(text="Fasting Blood Sugar")
```

***

# Rest ECG
```{r}
rest_ecg<-data2 %>% 
  group_by(restecg) %>% 
  summarise(count=n()) %>% 
  mutate(ecg_perc=paste0(round(count/sum(count)*100, 2), "%"))

view(rest_ecg)

highchart() %>% 
  hc_chart(type ="column",
           options3d = list(enabled = TRUE, alpha = 15, beta = 15)) %>%
  hc_xAxis(categories = rest_ecg$restecg) %>% 
  hc_add_series(data = rest_ecg$count, name = "Resting ECG") %>%
  hc_add_theme(hc_theme_bloom()) %>%
  hc_title(text="ECG")
```

***

# Chest pain
```{r}
chest_pain<-data2 %>% 
  group_by(cp) %>% 
  summarise(count=n()) %>% 
  mutate(cp_perc=paste0(round(count / sum(count) * 100, 1), "%"))


highchart() %>% 
  hc_chart(type ="bar",
           options3d = list(enabled = TRUE, alpha = 15, beta = 15)) %>%
  hc_xAxis(categories = chest_pain$cp) %>% 
  hc_add_series(data = chest_pain$count, name = "Chest Pain") %>%
  hc_add_theme(hc_theme_538()) %>%
  hc_title(text="Chest Pain Incidence")
```

***

# Age incidence
```{r}
age<-data2 %>% 
  group_by(age) %>% 
  count() %>% 
  filter(n>10)

highchart() %>% 
  hc_chart(type ="area",
           options3d = list(enabled = TRUE, alpha = 15, beta = 15)) %>%
  hc_xAxis(categories = age$age) %>% 
  hc_add_series(data = age$n, name = "Age Incidence") %>%
  hc_add_theme(hc_theme_google()) %>%
  hc_title(text="Age")

```

***

# Chest pain (cp) vs Blood Pressure (trtbps)
```{r}
data2 %>% 
  ggplot(aes(x=sex, y=trtbps))+
  geom_boxplot(fill='red')+
  facet_grid(~cp)+
  labs(title = "Corelation of Chest Pain with Blood Pressure", x="Sex", y="BP")

```

***

# Chest pain (cp) vs Cholesterol (chol)
```{r}
data2 %>% 
  ggplot(aes(x=sex, y=chol))+
  geom_boxplot(fill="orange")+
  facet_grid(~cp)+
  labs(title = "Corelation of Chest Pain with Cholesterol", x="Sex", y="Cholesterol")

```

# Chest pain in relation to Resting Blood Pressure With Age

```{r}
data2 %>% 
  ggplot(aes(x=age, y=trtbps, group=sex))+
  geom_boxplot(fill="purple")+
  facet_grid(~sex)+
  labs(title = "Corelation of Resting BP with Sex", x="Sex", y="TRTBP")


```


***

# Correlation
```{r}
data %>% 
  select(age, sex, cp, fbs, restecg, chol, exng, output, thall) %>% 
  cor() %>% 
  hchart()

```

# Conclusion

**Middle aged people are more prone to heart disease.**

  
**Age and Heart disease have positive correlation.**

  
**Cholesterol levels and Heart disease have positive correlation.**


**Men are prone to more cholesterol levels than women.**


**Exercise induced angina has negative impact on chest pain.**


**Women are more prone to heart disease than men.**


**Resting BP increases with Age and is more in Females than males.**

